---
title: 'Project FeederWatch: Zero-filling and taxonomic roll-up'
author: "Wesley Hochachka"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    number_sections: no
    theme: cerulean
    highlight: default
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Provide R code that enables users working with Project FeederWatch data to perform 
key data manipulations that are relevant to most valid uses of the data: 

 1. Zero-filling (i.e., adding counts of zero birds for species that were not reported
 as being detected during an observation period) 
 
 2. Taxonomic roll-up (i.e., combining subspecies and distinct taxonomic forms 
 when multiple species codes exist for focal taxa) 


# How to use this document

To use the R code in this document, you'll need to download the most up-to-date 
version of [R and RStudio](https://www.rstudio.com/products/rstudio/download/), 
and have some knowledge of the programming language R. 

To run the example code, you’ll need to download the sample checklist data from
the [Project FeederWatch website](https://feederwatch.org/explore/raw-dataset-requests/)
or the [GitHub repository](https://github.com/engagement-center/Project-FeederWatch-Zerofilling-Taxonomic-Rollup-Public). 

We also reference a "species translation table," which isn't needed to run the 
example code, but is important for determining which species code one needs. This 
table can be found on the [Project FeederWatch website](https://feederwatch.org/explore/raw-dataset-requests/). 

If you would like the most recent version of the complete FeederWatch dataset, it 
is also available on the [FeederWatch website](https://feederwatch.org/explore/raw-dataset-requests/). 
Note that the FeederWatch data are updated annually on or about June 1. The historic versions are
available in a [Mendeley Data archive](https://data.mendeley.com/datasets/cptx336tt5/2).


# Introduction

Project FeederWatch is a place-based program that asks participants to identify 
and count the birds that visit backyards, nature centers, community areas, and 
other locales in North America from November-April. Historically, the program
has focused on sites that provide supplementary feeding stations (i.e., bird feeders), 
but in the 2021-2022 season the project invited people to participate even if 
they didn't have a feeder. Moving forward participants are only required to have
an area with plantings, habitat, water, or food that attracts birds. 

At this time, the Project FeederWatch team recommends researchers exclude sites
that do not provide supplementary feeding stations until we can verify the best 
way to work with this new data.

The Project FeederWatch data is publicly available and distributed as .csv files 
compressed into ZIP archives. Those interested in using the data are strongly encouraged 
to consult the Project FeederWatch data report for tips on proper use of the data
([Bonter and Greig 2021](https://doi.org/10.3389/fevo.2021.619682)). 

For most uses, researchers will want to manipulate the data in one or two of the
following important ways before the data can be used for meaningful analyses:
 
 1. **Zero-filling:** Information on when birds were not reported as being detected
 should be added to the downloaded data (when appropriate) by inserting records 
 with counts of zero birds.
 
 2. **Taxonomic Roll-up:** The identities of some bird species may need to be modified
 because some of the reported information is being stored as subspecies or other
 recognizable forms (e.g., “Oregon” Junco instead of Dark-eyed Junco).
 
# Data structure

There are two data sets that make up the Project FeederWatch data: (1) checklists
(i.e., bird counts) and (2) site descriptions. This document focuses on the checklist
data, and does not go into detail on each data field in these data sets, the data 
collection protocols, or the data validation processes. To learn more, see the
[Data Dictionary available on the Project FeederWatch site](https://feederwatch.org/explore/raw-dataset-requests/)
and the published data report ([Bonter and Greig 2021](https://doi.org/10.3389/fevo.2021.619682)). 

The Project FeederWatch checklist data available for download do not contain counts
of zero, only observations of presence. More specifically, each row in the checklist
data features either the maximum number of each bird species seen in their count site
over a 2-day checklist period or a “nobird” entry to indicate that no birds were
seen on that 2-day checklist period. However, because the FeederWatch protocol
instructs participants to record all species seen within the count area, counts
of zero birds can be inferred and added to the available data. 

# What does "zero-filling" mean and why is it important?

Zero-filling is adding counts of zero birds for species that were not detected 
during an observation period. These zeros indicate either that a species was absent,
or that the species was present but the participant did not notice it. These two
alternatives need to be kept in mind when analyzing zero-filled data. 

For analyses, counts of zero are extremely important. For example, week-to-week 
changes in whether any birds of a species are seen at a site can provide 
information about birds’ motivation or need to obtain food from bird feeders. 
As another example, both the non-zero counts and the zero counts are needed to 
accurately describe changes in birds’ distribution and numbers through time. 
Without knowing which locations have counted zero birds of a species, there is 
no way of knowing whether the apparent geographic distribution of a species is a
description of the actual distribution or merely the distribution
of observers. Additionally, changes in numbers through time are a product both of
changes in the numbers of individuals at a site at which the species was reported,
as well as changes in the proportion of sites at which at least at least one
individual was seen. Both the non-zero counts and the zero counts are needed to
accurately describe changes in numbers through time. See Guillera-Arroita et al.
(2015) for a more detailed discussion of the importance of zero-count information.

As noted above, the double meaning of a zero—either as a true absence of a 
species, or merely the absence of detection of the species—does need to be taken
into account. There are two approaches that can be used to correctly interpret the zero counts.
One approach is to remove the ambiguity between true absence and lack of detection
by using statistical methods such as occupancy models (see Zuckerberg et al. 2010
for an example using data from Project FeederWatch) or N-mixture models (Goldstein 
et al. 2022). Because these two methods require data to be collected on repeated visits of 
the same set of locations, FeederWatch data are well suited for these methods. 

The second approach to correct interpretation of zero counts is to treat the zeros
as indicators of “relative occurrence”. This means one must assume that some 
constant proportion of the zero counts were at places and times at which the 
species was actually present but was undetected. The key is to analyze data in 
a way that tries to ensure that each zero has the same likelihood of meaning that
“the species was present but was not detected”. This is done by analyzing the data
in Generalized Linear Models in which predictors of variation in the probability
of detection have been included so that this variation is statistically removed.

With Project FeederWatch the major source of variation in detection rate of a species that we
have identified is the amount of time that an observer watches their site during
an observation period. This amount of observation effort is described by two 
variables in the FeederWatch data: the number of half-days of observation within
the two-day observation period, and the estimated number of effort-hours across
the entire two-day period. We have found that both of these variables should be
included as predictors whenever possible; however, only the first of these was
recorded in the earliest years of Project FeederWatch. See States et al. (2009)
for an example of how we have done this in our own analysis. 

## Important considerations when zero-filling

The process of zero-filling, as described here, is appropriate in any observational 
data, including Project FeederWatch data, under two conditions:
 
1. We have knowledge of observation periods during which the species of interest
was not seen.

2. It is reasonable to expect that the species of interest would have been reported
if it had been detected. 

For more information, refer to the section, "How to decide if zero-filling is appropriate?"
in this document where we have summarized the issues that one needs to consider
when wanting to zero-fill Project FeederWatch data for their species of interest.

# What does "taxonomic roll-up" mean and why is it important?

A taxonomic roll-up is combining subspecies and distinct taxonomic forms when multiple
species codes exist for focal taxa.

Project FeederWatch participants record the maximum numbers of individual birds
that they have seen at the same time during a two-day checklist period. For a subset
of species, participants are able to specify the level of recognizable form instead
of just reporting their observations at the taxonomic level of species. 

These recognizable forms are stored in the Cornell Lab of Ornithology database
using a “species code” that follows [eBird’s taxonomic system](https://ebird.org/science/use-ebird-data/the-ebird-taxonomy),
which is updated annually. We have created a species translation table following
this taxonomy that provides the species-level code for each recognizable form,
as well as the Latin, English, Canadian French, and Mexican Spanish names for each species. 

Examples of species with multiple recognizable forms include Dark-eyed Junco and
Fox Sparrow. It is possible for an observer to report some individual birds as
coming from a recognizable form and other individuals from the species without
identifying their form, and this can occur even during the same two-day checklist
period. Species like Rock Pigeon also have more than one value of SPECIES_CODE:
(1) the generic species, (2) truly wild Rock Pigeons inside their native range
(these are not found in data from Project FeederWatch), and (3) feral Rock Pigeons.
While most records of Rock Pigeons in the Project FeederWatch data are recorded 
under the generic species code of "rocpig", a non-trivial proportion of records
specify the feral form using the species code "rocpig1". 
 
Entering information on recognizable forms is optional in Project FeederWatch. 
To enter a recognizable form, a participant must enter information not provided 
on their standard “species to expect” list. As a result, one cannot assume that
the lack of a recognizable form at a site means that the recognizable form
was not present. The probability of a report of a recognizable form is some unknowable
joint probability of (1) a recognizable form being present at a feeder site and 
(2) the probability that an observer will be motivated enough to report that bird as a
recognizable form instead of as a member of a full species. This second probability,
participant motivation, will almost certainly vary among participants, across the
continent, and through time (because the ease of reporting recognizable forms has
increased through time). 

Thus, under most circumstances, if one wants to work with all of the data from 
such species, they will need to make sure that all data from a species will have
the same species code value. Following the jargon used for this same data manipulation
when working with data from eBird using the R package `auk`, we are referring to
this process as a "taxonomic roll-up."
 
## Important considerations when engaging in taxonomic roll-up

We have created code for a taxonomic roll-up that is relatively simple and assumes
Project FeederWatch participants are not double-counting the same individual 
birds both as belonging to a recognizable form and as members of the full species. 

Double-counting (the erroneous inflation of the number of individual birds being
reported for a species as a whole) could occur if a participant wanted to record
that they had seen some unusual individual from a recognizable form, but then 
also assume they needed to combine that unusual individual with the count of the
rest of the individual birds seen of that species. 

Hypothetical example: A participant who normally only sees Dark-eyed Juncos of the "Slate-colored"
form notices an "Oregon" form one day. The participant could report that "Oregon" junco
as one observation record, and additionally group the single unusual junco in with
all of the other juncos, reporting them all under the the species code for the full
species, "daejun." There may be ways of assessing whether this is happening frequently,
for example, by testing whether unexpectedly high counts of individual birds are
reported when both a recognizable form and the full species are reported during
a single observation period. 

The process of performing a taxonomic roll-up as described below could be modified
such that counts are only added together if all individuals are reported to be of
a recognizable form and there are no individuals recorded as coming from the full
species. An alternative approach could be to take the maximum number of individuals
reported for any of the recognizable forms and include that value as the number
observed for the species. 

Individual users of the data will need to decide how they want to handle the data
from recognizable forms. All users of Project FeederWatch data need to think
about this process, and include in the Methods section of any manuscript their
decisions and reasoning. Please refer to the section, "Additional
complexities with taxonomic roll-up" for a summary of issues one needs to 
consider about taxonomic roll-up.

# An R function for zero-filling and taxonomic roll-up

In this section we show the R code for creating a function that will both zero-fill 
data for one or more species, and perform a taxonomic roll-up. The function has been
set-up so that the default behavior will be to pass the data through the taxonomic 
roll-up process before the zero-filling process. However, depending on what is most
appropriate, the user has the ability to turn off the taxonomic roll-up process. 

At the end of this section we present examples of using this function. If you want 
to understand how this function works, there is a detailed explanation of how we
use R for the zero-filling and taxonomic roll-up processes in later sections of 
this document.

## Initial set-up in R

Load the `tidyverse` package. Also set the working directory to be where your 
copy of the data files are located.

```{r initial_setup, eval = TRUE, message=FALSE}
# Load the tidyverse package
library(tidyverse)

# Assign directory if not already done
# MyDirectory <- "D:/Project_FeederWatch_Analysis"
# Set working directory (i.e., where R looks for the data and stores output)
# setwd(MyDirectory)
```

Read in the sample checklist data.
```{r read_in_example_data, eval = TRUE}
PFW2021 <- read_csv("PFW_R_data_sample.zip")
```


## The generalized zero-filling and taxonomic roll-up function

Creating a function is, in the simplest cases, nothing more than bundling together
a set of R code so that all of the code can be run by simply typing the name that
is given to the function. The user will add "arguments" inside parentheses
immediately after the name of the function. Arguments to a function provide
a convenient way for users to specify changes to a small number of aspects of 
how the function works, for example specifying the table of data on which the function
needs to act. 

The function, named `ZeroFillPFW`, is defined using the command `function()`and 
accepts three arguments:
 
 1. `InputData`, the name of the data table (can be data.frame or a tibble) that will be used for zero-filling,
        
 2. `SpeciesCodes`, a vector of one or more SPECIES_CODE values for which zero-filling will be performed, and
 
 3. `rollup`, a logical (TRUE/FALSE) indicator of whether a taxonomic roll-up 
 should be done with the default behavior being to pass the data through the roll-up
 process before the zero-filling step.
 
Everything enclosed within the curly brackets is the set of R commands that will
be run when the function is run.

```{r declare_function, eval=TRUE}
ZeroFillPFW <- function(InputData, SpeciesCodes, rollup = TRUE){
  
  #The first three steps collect information that
  # will be necessary for taxonomic roll-up and zero-filling. First, we create
  # a table that contains only the information describing the observation 
  # periods (and not anything specific to any single birds species). Second,
  # we need to count the number of bird species for which zero-filling has been
  # requested, in order to know how many times we will need to loop through the
  # zero-filling process. Third, we need to check the species codes for which 
  # zero-filling was requested by comparing these to a list of all species codes 
  # found in InputData.
  
  #Create the table of sampling event information from all observation periods
  PFW_SED <- InputData %>%
    select(-OBS_ID, -SPECIES_CODE, -alt_full_spp_code, -HOW_MANY, -PLUS_CODE, 
           -VALID, -REVIEWED) %>%
    distinct()
  
  #Count the number of species codes
  NSpecies <- length(SpeciesCodes)
  
  #Identify any species codes that are not found in the data
  InvalidSppCodes <- InputData %>%
    select(SPECIES_CODE) %>%
    distinct() %>%
    anti_join(y = ., x = tibble(SPECIES_CODE = SpeciesCodes), by = "SPECIES_CODE")
  #For each non-existing species code, print an error message
  if (nrow(InvalidSppCodes) >= 1){
    for (i in 1:nrow(InvalidSppCodes)){
      cat(paste("THE SPECIES CODE \'", 
                InvalidSppCodes$SPECIES_CODE[i], 
                "\' DOES NOT EXIST\n",
                sep = ""))
      cat(" IN THE DATA TABLE, AND ZERO-FILLING IS NOT POSSIBLE.\n\n")
    }
    #break out of the function by returning a null result
    cat("THIS FUNCTION IS STOPPING BECAUSE ZERO-FILLING OF ALL\n")
    cat(" REQUESTED SPECIES CODES IS NOT POSSIBLE.\n\n")
    return(NULL)
  }
  
  #Loop through the species codes in turn, possibly doing a taxonomic roll-up,
  # and always zero-filling each species' data.
  for (i in 1:NSpecies){
    #
    #Check if this species code is "nobird" and if so skip over this 
    # species code.
    if (SpeciesCodes[i] == "nobird"){
      cat("THE SPECIES CODE \'nobird\' INDICATES AN ABSENCE OF BIRDS,\n")
      cat(" AND IS NOT AN ACTUAL BIRD SPECIES. NO ZERO-FILLING WILL\n")
      cat("BE DONE FOR THE SPECIES_CODE \'nobird\'.\n\n")
      next
    }
    #Place the SPECIES_CODE for the ith species into Spp_to_Zero_Fill.
    Spp_to_Zero_Fill <- SpeciesCodes[i]
    #
    #If a taxonomic roll-up is to be done, then do so for the focal species.
    if (rollup){
      InputData <- InputData %>%
        mutate(SPECIES_CODE = case_when(
          (is.na(alt_full_spp_code)
           ~ SPECIES_CODE),
          ((!is.na(alt_full_spp_code) &
              alt_full_spp_code == Spp_to_Zero_Fill)
           ~ alt_full_spp_code),
          ((!is.na(alt_full_spp_code) &
              alt_full_spp_code != Spp_to_Zero_Fill)
           ~ SPECIES_CODE))
        )
    }
    
    #Create template for zero-count records for the focal species.
    FakeSppSpecificFields <- as_tibble_row(list(OBS_ID = "OBSnull",
                                                SPECIES_CODE = Spp_to_Zero_Fill,
                                                HOW_MANY = 0,
                                                PLUS_CODE = NA,
                                                VALID = 1,
                                                REVIEWED = 0))
        
    #Zero-fill the data for the focal species.
    ZeroFilledFocalSpecies <- InputData %>%
      filter(SPECIES_CODE == Spp_to_Zero_Fill) %>%
      bind_rows(., bind_cols(PFW_SED, FakeSppSpecificFields)) %>%
      group_by(across(c(names(PFW_SED), "SPECIES_CODE"))) %>%
      summarize(OBS_ID = min(OBS_ID), HOW_MANY = sum(HOW_MANY), 
                PLUS_CODE = max(PLUS_CODE),
                VALID = min(VALID), REVIEWED = max(REVIEWED)) %>%
      ungroup()
    ZeroFilledFocalSpecies$PLUS_CODE <- as.logical(ZeroFilledFocalSpecies$PLUS_CODE)
    
    #Put the zero-filled data into a table that accumulates data from all of
    # the species named in the vector SpeciesCodes. Either create the tibble
    # to be returned if it does not already exist, or append rows to an
    # existing object.
    if ((exists("OutputData")))
      OutputData <- bind_rows(OutputData, ZeroFilledFocalSpecies)
    else
      OutputData <- ZeroFilledFocalSpecies
  }
  #After all species' data are zero-filled, return the zero-filled data to the 
  # user. Unless the user of the function directs the output into some object,
  # the output will be printed to the console and not be saved for subsequent use.
  #Also, if a taxonomic roll-up has been done, print a message to remind users
  # this was done.
  if (rollup){
    cat("\nNOTE: ANY RECORDS OF BIRDS REPORTED TO THE LEVEL OF A SUBSPECIES OR\n")
    cat(" OTHER RECOGNIZABLE FORM HAVE BEEN TREATED AS MEMBERS OF THE FULL\n")
    cat(" SPECIES FOR THE PURPOSE OF ZERO-FILLING, AND 'species_code' VALUES\n")
    cat(" FOR THESE RECOGNIZABLE FORMS HAVE BEEN REPLACE BY THE CODES FOR THE\n")
    cat(" FULL SPECIES IN THE OUTPUT.\n\n")
  }
  return(OutputData)
}
```

In addition to zero-filling and taxonomic roll-up code, the function is made robust to possible errors:

 - The function checks whether all of the SPECIES_CODE values actually exist in
 the data to be zero-filled; if not, then the function will not return any output. 

 - There is a check that "nobird" is not listed as a species for which zero-filled
 data are being requested. "nobird" is not a species of bird, but instead is an 
 indicator that no bird species was reported during a specific 2-day checklist period
 (i.e. for a specific SUB_ID). The "nobird" records exist in the downloaded data
 as a way of adding a zero-count record even when there were no bird species seen
 during the two-day checklist period. Otherwise, zero-count records erroneously
 would not exist for observation periods during which no bird species were recorded. 

 - A message is printed out whenever the "rollup" argument is set to "TRUE", and
 a taxonomic roll-up might have been done prior to zero-filling. 

## Examples of using the zero-filling and taxonomic roll-up function

Once the function has been defined by running the part of the R script that is 
inside the box immediately above, you’ll be able to use the function to run
through the examples below with sample checklist data from 2021.

### Example 1: Two-species taxonomic roll-up and zero-filling

Use the function `ZeroFillPFW` to perform a taxonomic roll-up and zero-fill data
for two species. Note that performing the taxonomic roll-up is the default and 
does not need to be specified.

The species referenced in this example are Rock Pigeon (rocpig) and Mourning Dove
(moudov). The species codes for these and all species can be found in the species
translation table mentioned at the beginning of this document and available for
download from the [Project FeederWatch website](https://feederwatch.org/explore/raw-dataset-requests/).

```{r zero_fill_eg_1, eval = FALSE, message = FALSE, results = "hide"}
MyZeroFilledData <- ZeroFillPFW(InputData = PFW2021, 
                                SpeciesCodes = c("rocpig", "moudov"))
```

### Example 2: Single-species zero-filling without taxonomic roll-up

The purpose of this example is to show how to tell the function not to do a taxonomic
roll-up. For the species used in this example, House Sparrow (species code: houspa), 
does not have any separate species code values for subspecies or other recognizable 
forms, so that the same results will be produced regardless of whether the taxonomic
roll-up is turned off or left on.

```{r zero_fill_eg_2, eval = FALSE, results = "hide"}
MyZeroFilledData_2 <- ZeroFillPFW(InputData = PFW2021, 
                                SpeciesCodes = "houspa", 
                                rollup = FALSE)
```

### Example 3: Potentially misleading zero-filling without a taxonomic roll-up

The purpose of this example is to compare the results of zero-filling with
and without taxonomic roll-up for a species that does have recognizable forms: the
Dark-eyed Junco (species code: daejun)

We first zero-fill with a taxonomic roll-up and then zero-fill a second time 
without the taxonomic roll-up. 

Finally, we count the number of non-zero records. The number of non-zero records
without the roll-up erroneously understates the number of checklist periods
during which our example species, Dark-eyed Junco (species code: daejun), was observed. 
This happens because none of the records using species codes from recognizable forms
are treated as Dark-eyed Juncos without the taxonomic roll-up.

```{r zero_fill_eg_3, eval = TRUE}
MyZeroFilledData_daejun_with_rollup <- ZeroFillPFW(InputData = PFW2021,
                                                   SpeciesCodes = "daejun",
                                                   rollup = TRUE)

MyZeroFilledData_daejun_without_rollup <- ZeroFillPFW(InputData = PFW2021,
                                                      SpeciesCodes = "daejun",
                                                      rollup = FALSE)

#Print to screen the counts on numbers of observation periods
# with and without taxonomic roll-up. 
MyZeroFilledData_daejun_with_rollup %>%
  filter(HOW_MANY > 0) %>%
  summarise(N_NonzeroRecords_with_rollup = n())

MyZeroFilledData_daejun_without_rollup %>%
  filter(HOW_MANY > 0) %>%
  summarise(N_NonzeroRecords_no_rollup = n())
```

### Example 4: Taxonomic roll-up to above the species level

For some purposes, it can be appropriate to combine information from multiple 
species together before zero-filling the data. One general case for doing this 
is for pairs or groups of species that cannot easily be distinguished from each other, 
for example Black-capped and Carolina Chickadees in their zones of geographic overlap and hybridization. 

Another example, considered here, are the hawks in the genus *Accipiter*. 
Accipiters are not easily identifiable to species, and in the FeederWatch data 
some sightings are even reported as being of an "Accipiter species" (species code ‘accipi’)
instead of being identified to species. If one wanted to examine patterns of geographic
or temporal change in perceived risk of depredation at sites, for instance, then it
may be reasonable to combine all records of any type of *Accipiter* into a single
taxonomic unit prior to zero-filling these data. This type of higher-than-species-level
taxonomic roll-up will need to be done prior to running the zero-filling function.

The first step in this process is to identify all of the possible values of SPECIES_CODE
that should be combined. This information can be found in the species-translation 
table. Searching this table for the text string "accipiter" shows that there are
8 SPECIES_CODE values for Accipiters (as of March 2022):
 
 + **shshaw**: the generic species code for Sharp-shinned Hawks
 + **shshaw2**: the code for the northern (i.e. North American) form of Sharp-shinned Hawk
 + **coohaw**: the generic species code for Cooper's Hawks
 + **y00612**: this species code indicates that the observer saw either a 
               Sharp-shinned or a Cooper's Hawk, but that the observer was not
               comfortable in identifying the bird to the level of a species
 + **norgos**: the generic species code for Northern Goshawk
 + **norgos2**: the code for the North American form of Northern Goshawk
 + **y00666**: this code indicates that the observer saw either a Cooper's 
               Hawk or a Northern Goshawk, but was not comfortable in making
               a species-level identification
 + **accipi**: the species code for "some type of *Accipiter* hawk"
 
We can replace all of the other species codes for *Accipiter* hawks with the code
"accipi" prior to zero-filling the data for the "accipi" species code. Below is 
R code that specifies how to replace the species codes before using the function
to zero-fill the data for this group of species.
 
```{r zero-fill_eg_4, eval = TRUE}
PFW2021_Accipiter_rollup <- PFW2021 %>%
  mutate(SPECIES_CODE = case_when(
    (SPECIES_CODE %in% c("shshaw", "shshaw2", "coohaw", "y00612",
                         "norgos", "norgos2", "y00666")
     ~ "accipi"),
    (TRUE
     ~ SPECIES_CODE)
  ))

MyZeroFilledData_4 <- ZeroFillPFW(InputData = PFW2021_Accipiter_rollup, 
                                SpeciesCodes = c("accipi"))
```

# How to decide if zero-filling is appropriate? 

Zero-filling is only appropriate if two conditions are met:

1. There is consistent information about the occurrence of observation periods 
during which the species of interest could have been seen although it was not actually
seen.

2. It is reasonable to assume that the species of interest would have been reported
as being seen if it had been detected.

## Condition 1 

The first condition is met by the Project FeederWatch data because of the nature 
of the project, with participants making observations from the same location multiple
times throughout the project’s season (November – April). There are records in the
database for each 2-day checklist period (typically a 2-day weekend), even if no birds were seen.

## Condition 2

The second condition may not always be valid because of the nature of how participants
report observations. Because participants are presented with a list of species, 
there are two classes of species in the data: (1) species that are on the lists 
that are presented to participants and should always be reported if seen (species
for which zero-filling can and should be done); and (2) write-in species that are
added by participants based on their own motivations (species for which zero-filling
may or may not make sense). 

Participants are able to add any species that are not on the list of expected species. 
Because adding these species requires additional effort from the participant, 
it is not safe to assume that all participants have gone through the effort 
of adding non-typical species to their lists. This is a generalization, but 
not a universal rule. 

For example, participants may be less inclined to report common species that are not typically 
seen at bird feeders, even though a feeder is not required to participate in the 
FeederWatch program as of the 2021-2022 season. Few participants, for instance, 
may be motivated enough to report species that at best could only marginally 
be considered to have visited a bird feeder (e.g., a Mallard on the pond beyond the bird feeders). 

Conversely, many participants would be thrilled to report a truly rare, vagrant 
species such as a Brambling (*Fringilla montifringilla*). As of March 2022 there
are eight locations that have records of Bramblings in the data of PFW. These are
likely an accurate record of all of the FeederWatch locations at which this rare
Eurasian vagrant has appeared. 

There is no record of which species are write-ins in which regions, and the lists 
of species presented to participants have changed through time for various reasons 
(e.g., a species has been expanding its range north). Additionally, these regions 
have become more geographically refined throughout the project’s history. Initially,
the continent was simply divided into “east” and “west”; online data entry has allowed
more precise lists to be generated. 

### How to determine is Condition 2 is met? 

To determine if the second condition (reasonable to assume that the species of 
interest would have been reported as being seen if it had been observed, for 
example if a species was a write-in species), we suggest that researchers use what
they know about the natural history of a species and information about how frequently
a species is reported when deciding if zero-filling is appropriate.

We have provided information about how frequently a species is reported in the 
species translation table. This table has a column labeled “n_locations”, and the
numbers in this column are counts of the number of unique FeederWatch sites at 
which each species has been reported at least once. Species reported at many sites
within their ranges are likely not write-in species and therefore good candidates
for zero-filling. For ubiquitous species like the Dark-eyed Junco, “many sites”
will mean tens of thousands of locations. For a range-restricted species in an 
area with a low density of FeederWatch participants, like Juniper Titmouse, a 
smaller number of locations may suffice as an indicator that the species is 
detected enough to warrant zero-filling.

## Final considerations regarding zero-filling

We are dealing with the issue of zero filling is such detail because we believe 
that it is essentially always beneficial, if not essential, to zero-fill Project
FeederWatch data before using them. An important reason for doing this (as already 
described) is that it is impossible to definitively detect changes in distribution
or abundance of a species through time, unless you can distinguish between the 
cases that: (1) an observer was present and would have reported a species if it
had been present, and (2) the absence of any data-collection activity. Only the
first of these alternatives provides useful information.

We are not aware of any circumstances under which more precise and accurate results
can be produced using statistical methods designed to work with “presence-only” 
(i.e. non-zero-filled) data. Such methods were developed for the modelling of 
species’ distributions when zero-count information are unavailable, which is the
case for records from museum collections. However, these methods are inherently 
limited in what they can tell us (Guillera-Arroita et al.2015). Further, these
presence-only analysis methods require the creation of “background” or “pseudo-absence” 
locations in lieu of actual locations at which zero individuals of a species were
reported. These “background” locations have no true observation event associated
with them, and so they do not have any information about observer effort (i.e. 
the amount of time that observers spent watching their site during an observation
period). As a consequence, the information about observer effort from the actual
records cannot be used in an analysis, even though we know that variation in 
observer effort accounts for substantial amounts of the variation in the number
of bird species observed and the number of individuals counted for any species 
(Bonter and Greig 2017). Failing to account for the effects of variable observer
effort will result in lower statistical power to detect actual patterns in the data.
For example, for a species distribution model, the distribution of a species will
not be as clearly defined (see Figure 2 of Johnston et al. 2021).

In summary, we strongly advise zero-filling data for the species for which this
is possible. If there is a species for which zero-filling is not possible and the
goal is to model the distributions of that species, then we suggest reading the 
following papers for guidance on appropriate analytical methods: Valavi et al. (2022)
and Valavi et al. (2021).

# Additional complexities with taxonomic roll-up

The complexities involved in taxonomic roll-ups have been mentioned earlier in 
this document. As illustrated in Example 3 above, taxonomic roll-up may be essential
for creating a data set that accurately describes when and where a species was reported. 
Conversely, the counts of the number of individual birds may be somewhat
inflated following a taxonomic roll-up. In practice, anyone who is using Project
FeederWatch data needs to consider the following three things about taxonomic roll-ups, 
and also to be able to justify in their methods the reason either for performing
(or not performing) a taxonomic roll-up:
 
 1. **A number of species will have some individual reports made to the level of a
 subspecies or other recognizable sub-group within that species.** For instance, 
 in Example 4 above, some participants have reported Sharp-shinned Hawks as belonging
 to the "northern" form, even though it is the only possible form within the FeederWatch
 region and potentially the observers did not even observe a bird closely enough to
 confirm without a doubt that the reported bird did indeed belong to the reported
 form. Basically, when using the data, always check whether the species in which
 you are interested has data reported from recognizable sub-specific forms. You
 can find whether some reports of your species of interest are at the sub-specific
 level by searching the species translation table.
 
 2. The converse can also happen, with **some observations being recorded as being
 identified only to the level above species.** *Accipiter* hawks, and Black-capped
 and Carolina Chickadees are two examples of this. Always check whether your species
 of interest is possibly ever or often reported as a member of some species pair
 or larger group (species codes for pairs of species are always the letter "y" 
 followed by 5 numbers). Very low frequencies of this are possibly ignorable noise. 
 If you need to treat your species of interest as part of a combined group, Example
 4 above provides an example of how to do so.

 3. **Some participants may be entering counts of recognizable sub-specific forms
 both as these forms, and then again as part of the total count for the overall 
 species.** A conservative approach would be to first check whether there are 
 observation periods in which your species of interest is simultaneously being 
 reported using the species code for the overall species and the species code for 
 some recognizable form. If this occurs, the most conservative course of action 
 would be to remove all of the data from the LOC_ID (i.e. the site) where this is
 found. For ubiquitous species like Dark-eyed Junco, the loss of data will be trivial. 
 Another option would be to assess whether the total counts are detectably higher
 during observation periods when both sub-specific forms and overall species are
 simultaneously reported; after this assessment, the data in question could be 
 retained or removed as appropriate.

# Step-by-step guide: Zero-filling for one species without taxonomic roll-up: 

If you are interested in seeing what is happening in the “black box” that is the
zero-filling and taxonomic roll-up function, continue reading for a heavily 
documented example of how to zero-fill the Project FeederWatch data for one
species without a taxonomic roll-up. 

This zero-filling process involves three main steps:

 1. Create a new row of data for every observation period, regardless of 
 whether the species of interest was seen during that observation period.
 
 2. Select a species to zero-fill and prepare a template for zero-count records. 
 
 3. Combine zero-count records created in step 1 and 2 with the original data. 
 This combining is done by summing together the counts of birds seen within each
 separate observation period. If there is an actual observation of that species
 being reported, the summed count is the actual count observed (any number plus
 zero is the original number). If the species was not reported in an observation
 period, then a new record of zero birds of that species is added to the table.
 
In order to run the R code in this section, download the sample checklist data 
from the [Project FeederWatch website](https://feederwatch.org/explore/raw-dataset-requests/)
or the [GitHub repository](https://github.com/engagement-center/Project-FeederWatch-Zerofilling-Taxonomic-Rollup-Public). Then, store the data in your working directory and read the data into R.

```{r read_in_example_data_1, eval = FALSE}
PFW2021 <- read_csv("PFW_R_data_sample.zip")
```

## Step 1: Create a new row of data for every 2-day checklist period

We need to create a basis for building the new records of counts of zero birds
for the bird species whose data are to be zero-filled. The basis for building these
zero-count records is the set of information from the actual observations that is
stored in the FeederWatch database, with each checklist period represented by 
a single row of information. 

The R commands below create the table of sampling event data, which are the subset
of variables shared across all species' records for a site-date combination.
This removes the variables that provide information that is unique to a single species,
and collapses these data down to a single unique copy of each record using the 
function `distinct()`.

```{r create_SED_1, eval = FALSE}
PFW_SED <- PFW2021 %>%
  select(-OBS_ID, -SPECIES_CODE, -HOW_MANY, -PLUS_CODE, 
         -VALID, -REVIEWED) %>%
  distinct()
```

## Step 2: Select species to zero-fill and prepare template for zero-count records

Zero-filling needs to be done on a species-by-species basis, because each species
will have a unique subset of 2-day checklist periods on which it was not observed. 
Once the species code for the species is assigned to an object in R, we can create
a template for inserting records that zero individuals of the species were observed. 
This template will then be appended to each row within the sampling event data 
that were created above. The template contains six pieces of information that 
are present for every observation of a species in the original data:
 
 1. **OBS_ID**: Each observation of a single species during a 2-day checklist period)
 is identified based on a unique OBS_ID value. Because we cannot assign a legitimate
 OBS_ID within the template, we will instead assign to the template a clearly 
 artificial value: "OBSnull". When combining a zero-count record with an existing 
 actual record in the second stage of zero-filling, “OBSnull" will be superseded 
 by the existing real OBS_ID.
 
 2. **SPECIES_CODE**: This is the code by which each taxon is referenced in the
 Project FeederWatch database.
 
 3. **HOW_MANY**: This is the maximum number of individual birds observed for 
 each species during each 2-day checklist period, and for zero-filling purposes,
 the appropriate number to use for the template is zero.
 
 4. **PLUS_CODE**: The very earliest Project FeederWatch data were submitted by
 participants on paper "bubble" forms on which the number of birds observed
 for a species needed to be indicated by filling in the bubble on the form that
 represented this specific number. Occasionally, the actual counts of birds that
 were seen was greater than the value of any of the bubbles. In these cases the
 participant would fill in a bubble that indicated a larger-than-maximum count,
 and the filling in of this bubble was reported as the value "1" in the column 
 labeled "PLUS_CODE". PLUS_CODE values are irrelevant for the zero-counts and
 this variable is assigned to be a missing value (represented by "NA").
 
 5. **VALID**: When an observation of a species is being sent into the database, 
 it passes through an automated filter that looks for unexpected species for a 
 given geographic area, or for unexpected high counts. If a record is not unusual,
 it is assigned a value of "1" in the variable VALID in the database. A "0" 
 indicates an unusual record that needs to be manually reviewed. If, upon review,
 the record appears to be legitimate, the value of VALID is changed to "1".
 Within the publicly available data, all records are available, regardless of their
 status based on the VALID variable. Because we want the zero-count records to
 remain if someone filters these data based on the VALID variable, we assign all
 of the new zero-count records a value of "1".
 
 6. **REVIEWED**: All records flagged as unexpected using the VALID variable 
 should eventually undergo a review by Project FeederWatch staff. When this review
 is complete, the presence of this manual review in denoted by setting the value 
 of the variable REVIEWED to be "1"; this is done, regardless of whether the 
 manual review led to any change in the VALID variable. The zero-count records 
 that we create are not manually reviewed, and so receive a value of "0".

```{r prep_single_spp_zero_filling, eval = FALSE}
#Specify the species code of the focal species for zero-filling.
Spp_to_Zero_Fill <- "daejun"

#Create fake zero-count records for a species of interest.
FakeSppSpecificFields <- as_tibble_row(list(OBS_ID = "OBSnull",
                                            SPECIES_CODE = Spp_to_Zero_Fill,
                                            HOW_MANY = 0,
                                            PLUS_CODE = NA,
                                            VALID = 1,
                                            REVIEWED = 0))
```

## Step 3: Do the zero-filling

Pipe the data through a series of manipulations that create the zero-filled final table of data. 

The following steps are taken in this process:
 
 1. Filter the input data down to retain only rows of input data from the focal 
 species that have been selected for zero-filling.
       
 2. Append to these real observations a set of fake observations in which there
 is a row for each observation period that is present in the input data, and in
 which each of these fake observations is a report of zero individuals of the focal species.
       
 3. Group the observations based on all of the names of variables in the `PFW_SED`
 tibble plus the species code for the focal species. After this, all summarizing
 operations will act within each of the groups, with each group either composed
 of two rows (an actual observation event that reported the focal species and a
 fake record of zero individuals being reported), or just the fake records of zero
 individuals for observation events during which no individuals of the focal 
 species were reported. We need to group based on all of the `PFW_SED` table's
 values plus SPECIES_CODE so that all of the values of these variables will
 appear in the output from the function `summarize()`.
       
 4. Collapse down the pairs of real and fake records so that:
 
     + the OBS_ID value is the actual OBS_ID when there is a pair of records,
               
     + the number of birds seen is the actual value when there is a real observation, or otherwise zero,
               
     + the PLUS_CODE is either the real value of PLUS_CODE of otherwise it is a missing value,
               
     + the VALID flag is set to the true valid flag when a real observation
               is present, or otherwise is set to a value of "1" (i.e. a valid
               record), and
               
     + the REVIEWED flag is set to the actual value when a real observation
               exists, or otherwise is set to "0" (i.e. not reviewed).
               
 5. Ungroup (remove the grouping information) so that further work with the 
 zero-filled data will not produce weird results. 
       
 6. Set the data type of the variable PLUS_CODE back to being a Boolean ("logical")
 data type within R, to match the data type of the input data. Somehow the data
 type has been changed in the process of running through the five steps in 
 the sequence of piped instructions.
 
```{r zero_filling_1, eval = FALSE}
ZeroFilledFocalSpecies <- PFW2021 %>%
  filter(SPECIES_CODE == Spp_to_Zero_Fill) %>%
  bind_rows(., bind_cols(PFW_SED, FakeSppSpecificFields)) %>%
  group_by(across(c(names(PFW_SED), "SPECIES_CODE"))) %>%
  summarize(OBS_ID = min(OBS_ID), HOW_MANY = sum(HOW_MANY), 
                         PLUS_CODE = max(PLUS_CODE),
                VALID = min(VALID), REVIEWED = max(REVIEWED)) %>%
  ungroup() %>%
  select(LOC_ID, LATITUDE, LONGITUDE, housing_density, SUBNATIONAL1_CODE, 
         ENTRY_TECHNIQUE, SUB_ID, OBS_ID, Month, Day, Year, PROJ_PERIOD_ID, 
         SPECIES_CODE, HOW_MANY, PLUS_CODE, VALID, REVIEWED, DAY1_AM, DAY1_PM, 
         DAY2_AM, DAY2_PM, EFFORT_HRS_ATLEAST, SNOW_DEP_ATLEAST, 
         Data_Entry_Method)
ZeroFilledFocalSpecies$PLUS_CODE <- as.logical(ZeroFilledFocalSpecies$PLUS_CODE)
```

Et voila! There is now a table named `ZeroFilledFocalSpecies` that contains 
the zero-filled records for the focal species.

# Step-by-step guide: Zero-filling for one species with taxonomy roll-up: 

The process of zero-filling data is the logical place in the sequence of data 
processing to conduct taxonomic roll-up. This is because the roll-up process 
needs to take into account the possibility that multiple reports of the same 
species may exist for an observation event, in the case in which some individuals
are reported as recognizable forms and other individuals are reported at the 
taxonomic level of the full species. Summing the counts of individuals within 
each full species and observation period is one possible approach.

Zero-filling is already summing counts of individuals from at least two rows for
each species and count period when a species was present: an additional 
dummy record with a count of zero individual birds needs to be summed within that
species and count period. This summing process, as implemented in the first example
of zero-filling without taxonomic roll-up, will already automatically add together
all of the two (or more) records that it finds that have the same SPECIES_CODE
and SUB_ID (i.e., the zero-filling process is pre-adapted to additionally aid
in the process of a taxonomic roll-up).

First, read in the data if you haven’t already.

```{r read_in_example_data_2, eval = FALSE}
PFW2021 <- read_csv("PFW_R_data_sample.zip")
```

## Step 1. Create a new row of data for every single observation period

Create a table that contains only the data describing the observation period 
(i.e., 2-day checklist) by removing all species-specific columns. Then reduce
this table down to only a single record for each observation period using `distinct()`.

```{r create_SED_2, eval = FALSE}
PFW_SED <- PFW2021 %>%
  select(-OBS_ID, -SPECIES_CODE, -alt_full_spp_code, -HOW_MANY, -PLUS_CODE, 
         -VALID, -REVIEWED) %>%
  distinct()

```

## Step 2: Replace sub-specific taxon codes for the focal species code

Roll up the records for the species whose data are to be zero-filled by replacing
the species code(s) for the recognizable form with the species code for the full 
species. The result of this replacement is put into a new tibble so that the data
read into R is free from modification in case it's needed for some other purpose
later in this R session. 

We use the `case_when()` function to do the replacing only for the single species
for which we want to zero-fill the data. There are three possible cases we need to
deal with: 
 
 1. If there is no full-species code in the field alt_full_spp_code then the 
 current value in SPECIES_CODE is retained.
     
 2. If there is a full-species code in the field alt_full_spp_code and that 
 value **is** for the species in which we are interested, then the value in 
 SPECIES_CODE is replaced by the value in alt_full_spp_code.
     
 3. If there is a full-species code in the field alt_full_spp_code and that
 value is not for the species in which we are interested, then the current 
 value in SPECIES_CODE is retained.

```{r roll_up_spp_codes, eval = FALSE}
#Specify the species code of the focal species for zero-filling.
Spp_to_Zero_Fill <- "rocpig"

PFW2021_rollup <- PFW2021 %>%
  mutate(SPECIES_CODE = case_when(
    (is.na(alt_full_spp_code)
      ~ SPECIES_CODE),
    ((!is.na(alt_full_spp_code) &
      alt_full_spp_code == Spp_to_Zero_Fill)
      ~ alt_full_spp_code),
    ((!is.na(alt_full_spp_code) &
       alt_full_spp_code != Spp_to_Zero_Fill)
     ~ SPECIES_CODE))
    )
```

## Step 3: Prepare template for zero-count records 

```{r zero_filling_2, eval = FALSE}
#Create fake zero-count records for the specified focal species.
FakeSppSpecificFields <- as_tibble_row(list(OBS_ID = "OBSnull",
                                            SPECIES_CODE = Spp_to_Zero_Fill,
                                            HOW_MANY = 0,
                                            PLUS_CODE = NA,
                                            VALID = 1,
                                            REVIEWED = 0))
```

## Step 4: Do the zero-filling

Pipe the data through a series of manipulations that create the zero-filled final
table of data. These are the same steps as in the first example in which there
was no taxonomic roll-up. 

```{r zero_filling_3, eval = FALSE}
ZeroFilledFocalSpecies <- PFW2021_rollup %>%
  filter(SPECIES_CODE == Spp_to_Zero_Fill) %>%
  bind_rows(., bind_cols(PFW_SED, FakeSppSpecificFields)) %>%
  group_by(across(c(names(PFW_SED), "SPECIES_CODE"))) %>%
  summarize(OBS_ID = min(OBS_ID), HOW_MANY = sum(HOW_MANY), 
            PLUS_CODE = max(PLUS_CODE),
            VALID = min(VALID), REVIEWED = max(REVIEWED)) %>%
  ungroup() %>%
  select(LOC_ID, LATITUDE, LONGITUDE, housing_density, SUBNATIONAL1_CODE, 
         ENTRY_TECHNIQUE, SUB_ID, OBS_ID, Month, Day, Year, PROJ_PERIOD_ID, 
         SPECIES_CODE, HOW_MANY, PLUS_CODE, VALID, REVIEWED, DAY1_AM, DAY1_PM, 
         DAY2_AM, DAY2_PM, EFFORT_HRS_ATLEAST, SNOW_DEP_ATLEAST, 
         Data_Entry_Method)
ZeroFilledFocalSpecies$PLUS_CODE <- as.logical(ZeroFilledFocalSpecies$PLUS_CODE)
```

The only difference between zero-filling with and without a taxonomic roll-up
is that there is an extra step before the zero-filling in which the SPECIES_CODE
values for recognizable forms are replaced by the SPECIES_CODE of the associated
full species.
 
Replacing existing SPECIES_CODE values with new species codes that aggregate together
taxa creates the possibility of doing other sorts of taxonomic roll-ups, such as
combining data from all *Accipiter* species into a single taxonomic unit by manually
replacing all of the species-level or lower-level species code values with some 
identical above-species-level species code. 

This might actually be a useful process for many uses (e.g. quantifying frequencies
of predators' visits to sites) because these *Accipiter* hawk species can be 
challenging to identify to species. Note that in the case of *Accipiter* hawks 
there is already a *SPECIES_CODE* for this, "accipi", and additionally there are
*species_code* values for "Sharp-shinned or Cooper's Hawk", and "Cooper's Hawk 
or Goshawk". All of these exist within the FeederWatch data.
 
# References

Bonter, D.N. and E.I. Greig. 2021. Over 30 years of standardized bird counts at 
supplementary feeding stations in North America: A citizen science data report 
for Project FeederWatch. Frontiers in Ecology and Evolution 9: Article 619682 [doi.org/10.3389/fevo.2021.619682](doi.org/10.3389/fevo.2021.619682) 

Goldstein, B. R., and P. de Valpine. 2022. Comparing N-mixture models and GLMMs
for relative abundance estimation in a citizen science dataset. Scientific Reports
12:12276. [doi.org/10.1038/s41598-022-16368-z](doi.org/10.1038/s41598-022-16368-z)

Guillera-Arroita, G., J. J. Lahoz-Monfort, J. Elith, A. Gordon, H. Kujala, P. E.
Lentini, M. A. McCarthy, R. Tingley, and B. A. Wintle. 2015. Is my species distribution
model fit for purpose? Matching data and models to applications. Global Ecology
and Biogeography 24:276-292. [doi.org/10.1111/geb.12268](doi.org/10.1111/geb.12268)

Johnston, A., W. M. Hochachka, M. Stimas-Mackey, V. Ruiz-Gutierrez, O. J. Robinson,
E. T. Miller, T. Auer, S. T. Kelling, and D. Fink. 2021. Analytical guidelines to
increase the value of eBird data to estimate species occurrence. Diversity and 
Distributions 27:1265-1277. [doi.org/10.1111/ddi.13271](doi.org/10.1111/ddi.13271)

States, S. L., W. M. Hochachka, and A. A. Dhondt. 2009. Spatial variation in an 
avian host community: implications for disease dynamics. Ecohealth 6:540-545. 
[doi.org/10.1007/s10393-009-0269-2](doi.org/10.1007/s10393-009-0269-2)

Valavi, R., G. Guillera-Arroita, J. J. Lahoz-Monfort, and J. Elith. 2022. 
Predictive performance of presence-only species distribution models: a benchmark
study with reproducible code. Ecological Monographs 92:e01486. 
[doi.org/10.1002/ecm.1486](doi.org/10.1002/ecm.1486)

Valavi, R., J. Elith, J. J. Lahoz-Monfort, and G. Guillera-Arroita. 2021. 
Modelling species presence-only data with random forests. Ecography 44:1731-1742.
[doi.org/10.1111/ecog.05615](doi.org/10.1111/ecog.05615)

Zuckerberg, B., D. N. Bonter, W. M. Hochachka, W. D. Koenig, A. T. DeGaetano, and 
J. L. Dickinson. Climatic constraints on wintering bird distributions are modified by 
urbanization and weather. 2010. Journal of Animal Ecology, 80:403-413.
[doi.org/10.1111/j.1365-2656.2010.01780.x](doi.org/10.1111/j.1365-2656.2010.01780.x)

# Acknowledgements

Sincere thank you to the thousands of participants who are part of the Project FeederWatch
community. Your dedication to recording the birds you see in your yards, and the 
fees you pay every season are why FeederWatch has been able to run for over 30 
years.

Thank you to Dave Bonter, Emma Greig, and Rachael Mady for their feedback and 
revisions. Thank you as well to Christopher Sayers and Laura Vander Meiden for reviewing 
this documentation. 

# Have feedback? 

This is the first documentation we have put together that includes R code and 
accompanying guidance for working with the FeederWatch data. Please let us know 
if there is anything missing or ways we could improve this documentation by contacting
Project FeederWatch via email (feederwatch@cornell.edu).
